Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-03-31 PAPER.tex
added conclusion paragraph to jackknife section

We find that the baselines test is fully consistent with noise, as all error bars pass through zero. Therefore, we conclude that particular baselines are not to be blamed for the bias in our dataset. For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. This indicates a low level bias that is LST-dependent, and likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing. The even/odd test shows $81\%$ of all error bars being consistent with zero, again indicating the presence of a low level bias. The cause of this particular systematic is not obvious, and requires further investigation that we do not carry out in this paper. Most importantly, a clean detection of EoR would have passed all three tests, which is clearly not what we see.

@@

There are a couple possible explanations for this. Although we expect each cross-multiplication that is used in power spectrum estimation to have independent noise, there is still the possibility of a noise-foreground coupling term that can introduce power. This is because cross-multiplications produce four additive terms --- a signal-squared term (where signal includes both foregrounds and EoR), two cross-terms between the signal and noise, and one noise-only term. When differencing two power spectra, we expect the signal-only term to subtract out for a `passing' null test, and we expect the noise-only terms to be consistent with thermal noise. The cross-terms, however, can introduce extra variance that would vary with foreground strength, therefore being consistent with the shape of our results. That being said, an alternate explanation for the large error bars seen at high $k$'s could simply be that our estimate of the thermal noise is too low, as all three tests seem to imply a larger, but consistent, estimate.

>>>>

We find that the baselines test is fully consistent with noise, as all error bars pass through zero. Therefore, we conclude that the baselines making up the two jackknives either do not contain bias, or contain similar amounts of bias. We suspect it is the former, though more thorough jackknives along this axis are needed to make this conclusion. For the LST test, $90\%$ of all the error bars pass through zero, which is slightly lower than the expected $95\%$. This indicates a low level bias that is LST-dependent, and likely caused by residual foregrounds that vary in LST. In other words, the two jackknife halves differ by an amount greater than thermal noise and do not completely cancel out the sky signal when differencing. The even/odd test shows $81\%$ of all error bars being consistent with zero, again indicating the presence of a low level bias. The cause of this particular systematic is not obvious, and requires further investigation that we do not carry out in this paper. Most importantly, a clean detection of EoR would have passed all three tests, which is clearly not what we see.

@@

There are a couple possible explanations for this. Although we expect each cross-multiplication that is used in power spectrum estimation to have independent noise, there is still the possibility of a noise-foreground coupling term that can introduce power. This is because cross-multiplications produce four additive terms --- a signal-squared term (where signal includes both foregrounds and EoR), two cross-terms between the signal and noise, and one noise-only term. When differencing two power spectra, we expect the signal-only term to subtract out for a ``passing" null test, and we expect the noise-only terms to be consistent with thermal noise. The cross-terms, however, can introduce extra variance that would vary with foreground strength, therefore being consistent with the shape of our results.  One additional explanation, specifically for the slightly large error bars seen at high $k$'s, could simply be that our estimate of the thermal noise is too low, as all three tests seem to imply a larger, but consistent, estimate in this high-$k$ regime. Regardless of a potentially larger thermal noise prediction, however, the large error bars near the horizon limit still signify the presence of excess variance in our data.

In this section we have presented the first jackknife and null tests from the PAPER experiment. While simple, they outline a framework that can be used in the future. The 21\,cm community is beginning to recognize the importance of these types of tests (\citealt{pober_et_al2016b}) in characterizing power spectra at the EoR level, and it is clear that future results will require more substantial and thorough tests.
